---
title: "Assignment2"
author: "Joseph Finke"
date: "12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction


## Wrangling, & Exploring Data

```{r loadings}
#Setting Work Directory
setwd("C:/Users/Josep/OneDrive/School/Fall 2021/D590 - Applied Data Science/Project")

#Loading Libraries
library(tidyverse)
library(dplyr)

#Loading the data files
Knowledge <- read.csv("KnowledgeData.csv", header=T)
Skills <- read.csv("SkillsData.csv", header=T)
Abilities <- read.csv("AbilitiesData.csv", header=T)

#Checking out the data
dim(Knowledge)
dim(Skills)
dim(Abilities)

# Adding datasets vertically
KSAdata <- rbind(Knowledge, Skills, Abilities)

#Checking out data
dim(KSAdata)
glimpse(KSAdata)

#View data
View(KSAdata)
str(KSAdata)

#Number of Occupations in dataset
nrow(table(KSAdata$Title))

#Number of KSAs by type
table(KSAdata$Element_Type)
```
## Visualizations

```{r visualizations}
#load library
library(ggplot2)

#Use skills data to example.
Skills2 <- read.csv("SkillsData2.csv", header=T)

#Compare occupations on one skill - active learning
ggplot(data=Skills2, aes(jobID, Active.Learning)) + geom_bar(stat = "identity", fill = "darkblue")

#Similiar, only horizontal
barplot(Skills2$Active.Learning, horiz = TRUE)

#Scatterplot of active learning & active listening
plot(Skills2$Active.Learning, Skills2$Active.Listening)

```

```{r visualization future}
#Flip rows and columns to better explore occupation's skills
Skills3 <- read.csv("SkillsData3.csv", header=T)

#Example plot of the skills needed by Chief Executives
ggplot(data=Skills3, aes(Skill, Chief.Executives)) + geom_bar(stat = "identity", fill = "darkblue") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

#Example plot of the skills needed by Transporation Security Screeners
ggplot(data=Skills3, aes(Skill, Transportation.Security.Screeners)) + geom_bar(stat = "identity", fill = "darkblue") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


## Clustering


### Measuring Distance
In this first subsection, I compute the distance or similiarity between each occupation's skill scores. I use the default Euclidean distance.

```{r distance}
#load libraries
library(cluster)
library(factoextra)

#set dataframe
df <- Skills2[,2:36]
df <- scale(df)

#compute disatnce matrix
distance <- get_dist(df)

#visualize distance matrix
fviz_dist(distance, gradient = list(low = "#0b5394", mid = "white", high = "#ef791a"))
```
### Number of K Clusters

Next, I attempt to find the optimal number of clusters using the eblow method. 

For this subsection, I adapted code from R-bloggers.

Citation:
* Title: Finding Optimal Number of Clusters
* Author: Sunny Anand
* Date: February 2017
* Availability: https://www.r-bloggers.com/2017/02/finding-optimal-number-of-clusters/

```{r elbow method}
#Elbow Method for finding the optimal number of clusters
set.seed(123)

# Compute and plot wss for k = 2 to k = 15.
k.max <- 15
wss <- sapply(1:k.max, 
              function(k){kmeans(df, k, nstart=50,iter.max = 15 )$tot.withinss})
wss
plot(1:k.max, wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")
```

### K Means Clustering & Visualizing Clusters
Next, I conducted k-means clustering. I also visualize the clusters. 

Portions of the following code were adapted from the University of Cincinnati Business Analytics R Programming Guide.
Citation:
* Title: UC Business Analytics R Programming Guide
* Author: Bradley Boehmke
* Availability: https://uc-r.github.io/kmeans_clustering

```{r kmeans}

#Build Clustering
k2 <- kmeans(df, centers = 2, nstart = 25)
k3 <- kmeans(df, centers = 3, nstart = 25)
k4 <- kmeans(df, centers = 4, nstart = 25)
k5 <- kmeans(df, centers = 5, nstart = 25)
k6 <- kmeans(df, centers = 6, nstart = 25)

#plots to compare
p1 <- fviz_cluster(k2, geom = "point", data = df) + ggtitle("k = 2")
p2 <- fviz_cluster(k3, geom = "point",  data = df) + ggtitle("k = 3")
p3 <- fviz_cluster(k4, geom = "point",  data = df) + ggtitle("k = 4")
p4 <- fviz_cluster(k5, geom = "point",  data = df) + ggtitle("k = 5")
p5 <- fviz_cluster(k5, geom = "point",  data = df) + ggtitle("k = 6")

library(gridExtra)
grid.arrange(p1, p2, p3, p4, p5, nrow = 2)
```

### Hierarchical Clustering
```{r}
#Loading the data files
topskill <- read.csv("SkillsData9.csv", header=T)

str(topskill)

# Compute distances
distances <- dist(topskill[2:36], method = "euclidean")

# Hierarchical clustering
clustertopskill <- hclust(distances, method = "ward")

#ClusterGroups
clusterGroups = cutree(clustertopskill, k = 25)

clusterGroups

# Plot the dendrogram
plot(clusterGroups)

cluster1 = subset(topskill, clusterGroups == 1)
# Look at the first 10 titles in this cluster:
cluster1$jobID[1:10]

cluster2 = subset(topskill, clusterGroups == 2)
# Look at the first 10 titles in this cluster:
cluster2$jobID[1:10]

cluster3 = subset(topskill, clusterGroups == 3)
# Look at the first 10 titles in this cluster:
cluster3$jobID[1:10]

cluster4 = subset(topskill, clusterGroups == 4)
# Look at the first 10 titles in this cluster:
cluster4$jobID[1:10]

cluster5 = subset(topskill, clusterGroups == 5)
# Look at the first 10 titles in this cluster:
cluster5$jobID[1:10]

cluster6 = subset(topskill, clusterGroups == 6)
# Look at the first 10 titles in this cluster:
cluster6$jobID[1:10]

cluster7 = subset(topskill, clusterGroups == 7)
# Look at the first 10 titles in this cluster:
cluster7$jobID[1:10]

cluster8 = subset(topskill, clusterGroups == 8)
# Look at the first 10 titles in this cluster:
cluster8$jobID[1:10]

cluster9 = subset(topskill, clusterGroups == 9)
# Look at the first 10 titles in this cluster:
cluster9$jobID[1:10]

cluster10 = subset(topskill, clusterGroups == 10)
# Look at the first 10 titles in this cluster:
cluster10$jobID[1:10]

cluster11 = subset(topskill, clusterGroups == 11)
# Look at the first 10 titles in this cluster:
cluster11$jobID[1:10]

cluster12 = subset(topskill, clusterGroups == 12)
# Look at the first 10 titles in this cluster:
cluster12$jobID[1:10]

cluster13 = subset(topskill, clusterGroups == 13)
# Look at the first 10 titles in this cluster:
cluster13$jobID[1:10]

cluster14 = subset(topskill, clusterGroups == 14)
# Look at the first 10 titles in this cluster:
cluster14$jobID[1:10]

cluster15 = subset(topskill, clusterGroups == 15)
# Look at the first 10 titles in this cluster:
cluster15$jobID[1:10]

cluster16 = subset(topskill, clusterGroups == 16)
# Look at the first 10 titles in this cluster:
cluster16$jobID[1:10]

cluster17 = subset(topskill, clusterGroups == 17)
# Look at the first 10 titles in this cluster:
cluster17$jobID[1:10]

cluster18 = subset(topskill, clusterGroups == 18)
# Look at the first 10 titles in this cluster:
cluster18$jobID[1:10]

cluster19 = subset(topskill, clusterGroups == 19)
# Look at the first 10 titles in this cluster:
cluster19$jobID[1:10]

cluster20 = subset(topskill, clusterGroups == 20)
# Look at the first 10 titles in this cluster:
cluster20$jobID[1:10]

cluster21 = subset(topskill, clusterGroups == 21)
# Look at the first 10 titles in this cluster:
cluster21$jobID[1:10]

cluster22 = subset(topskill, clusterGroups == 22)
# Look at the first 10 titles in this cluster:
cluster22$jobID[1:10]

cluster23 = subset(topskill, clusterGroups == 23)
# Look at the first 10 titles in this cluster:
cluster23$jobID[1:10]

cluster24 = subset(topskill, clusterGroups == 24)
# Look at the first 10 titles in this cluster:
cluster24$jobID[1:10]

cluster25 = subset(topskill, clusterGroups == 25)
# Look at the first 10 titles in this cluster:
cluster25$jobID[1:10]


topskill$cluster <- clusterGroups


cluster25$jobID[1:10]
```


